---
title: "CodeBook"
author: "Chris Bowman"
date: "November 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Peer-graded Assignment: Getting and Cleaning Data Course Project
Set data directory for my data
and verify that there is or is not a ./data directory. 
 Name and download the file to be utilized.

```{r}
setwd("C:/R_Projects")

if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
```
Download.file(fileUrl,destfile="./data/Dataset.zip",method="wininet") and then unzip the dataset
```{r}
unzip(zipfile="./data/Dataset.zip",exdir="./data")
```
Name the file path, list all files, and display them
```{r}
path_rf <- file.path("./data" , "UCI HAR Dataset")
files<-list.files(path_rf, recursive=TRUE)
files
```

 Begin the process of reading the activity files related to the training set and the test set.  
 Begin the process of reading the subject files related to the training set and the test set.  
 Begin the process of reading the features files related to the training set and the test set.
```{r}
dataActivityTest  <- read.table(file.path(path_rf, "test" , "Y_test.txt" ),header = FALSE)
dataActivityTrain <- read.table(file.path(path_rf, "train", "Y_train.txt"),header = FALSE)

dataSubjectTrain <- read.table(file.path(path_rf, "train", "subject_train.txt"),header = FALSE)
dataSubjectTest  <- read.table(file.path(path_rf, "test" , "subject_test.txt"),header = FALSE)

dataFeaturesTest  <- read.table(file.path(path_rf, "test" , "X_test.txt" ),header = FALSE)
dataFeaturesTrain <- read.table(file.path(path_rf, "train", "X_train.txt"),header = FALSE)
```
View the data and verify that data was read to each table (optional)

dataActivityTest 

dataActivityTrain

dataSubjectTest

dataSubjectTrain

dataFeaturesTest

dataFeaturesTrain

Display the structure of each object
```{r}
str(dataActivityTest)
str(dataActivityTrain)

str(dataSubjectTest)
str(dataSubjectTrain)

str(dataFeaturesTest)
str(dataFeaturesTrain)
```
This develops an object by merging the training and the test sets to create one data set. Each new dataset is developed by using two files differentiated as to "train" and "test." The new table is made by binding the rows for both files to create a new object for subject, activity, and features.
```{r}
dataSubject <- rbind(dataSubjectTrain, dataSubjectTest)
dataActivity<- rbind(dataActivityTrain, dataActivityTest)
dataFeatures<- rbind(dataFeaturesTrain, dataFeaturesTest)
```
This adds the names to the variables
```{r}
names(dataSubject)<-c("subject")
names(dataActivity)<- c("activity")
dataFeaturesNames <- read.table(file.path(path_rf, "features.txt"),head=FALSE)
names(dataFeatures)<- dataFeaturesNames$V2
```
A new data frame, datacombine, is developed by binding all columns from the three distinct files to get the final data frame Data for analysis.
```{r}
dataCombine <- cbind(dataSubject, dataActivity)
Data <- cbind(dataFeatures, dataCombine)
```

This makes a new object by selecting out those observation marked by the terms "mean" amd "std" to make a new subset FeaturesNames by measurements on the mean and standard deviation.
```{r}
subdataFeaturesNames<-dataFeaturesNames$V2[grep("mean\\(\\)|std\\(\\)", dataFeaturesNames$V2)]
```
Subset the data frame, Data, by selected names of Features
```{r}
selectedNames<-c(as.character(subdataFeaturesNames), "subject", "activity" )
Data<-subset(Data,select=selectedNames)
```
Check the structures of the data frame Data
```{r}
str(Data)
```
This uses descriptive activity names from the file activity_labels.txt to name the activities in the data set so as to provide clarity.
```{r}
activityLabels <- read.table(file.path(path_rf, "activity_labels.txt"),header = FALSE)

head(Data$activity,30)
```
Appropriately labels the data set with descriptive variable names. Variables activity and subject and names of the activities have been labelled using descriptive names. In this part, Names of Features labels variable names.

Prefix t is replaced by time, Acc is replaced by Accelerometer, Gyro is replaced by Gyroscope, prefix f is replaced by frequency, Mag is replaced by Magnitude, and BodyBody is replaced by Body
```{r}
names(Data)<-gsub("^t", "time", names(Data))
names(Data)<-gsub("^f", "frequency", names(Data))
names(Data)<-gsub("Acc", "Accelerometer", names(Data))
names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
names(Data)<-gsub("Mag", "Magnitude", names(Data))
names(Data)<-gsub("BodyBody", "Body", names(Data))

names(Data)
```

This builds a text file, tidydata.txt, set and saves it.
```{r}
library(plyr);
Data_Aggr<-aggregate(. ~subject + activity, Data, mean)
Data_Aggr<-Data_Aggr[order(Data_Aggr$subject,Data_Aggr$activity),]
setwd("C:/R_Projects")
write.table(Data_Aggr, file = "./tidydata.txt",row.name=FALSE)
```

